---
title: "T2DM meta analysis"
author: "ks"
date: "3/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# {.tabset}

## load Libraries and Data

```{r packages}
for (pkg in c("tidyverse","MASS","purrr", "readxl","dplyr","meta","metafor","netmeta")) 
   library(pkg, character.only = TRUE)

str_detect = stringr::str_detect
```


```{r load-data}
data.loc <- c("~kims/Google Drive/My Drive/Teaching/PM605/pm-605-sp2021/week5-feb22/")
dmbc <-  paste0(data.loc, 
                                      "PM605DiabetesAndBlad_DATA_2019-05-03_1706_CLEANED.rds") %>% 
  readRDS()
```

```{r view}
dim(dmbc)
head(dmbc)
```


```{r study-design}
dmbc %>% group_by(study_design) %>% tally()
```

```{r prep-data}

ma.dat <-  dmbc %>% 
              mutate(
                lnes = log(ratio_est),
                lnes_l = log(ratio_est_l95),
                lnes_u = log(ratio_est_u95),
                selnes = (lnes_u-lnes_l)/3.919928,
                Study = paste0(first_auth," et al., " , datepub),
                cases = affected_exp + affected_unexp,
    
        # order design variable show it shows correctly in the plot
                Design = study_design %>% forcats::fct_relevel(c("1. Pro Cht","2. Pop CC","3. Ret Cht",
                                                           "4. SIR","5. Hosp CC"))
  )

```

## Meta-analysis

Fit a random effects meta analysis to all the data.

```{r metaa}
rema <- metagen(lnes, selnes, data = ma.dat, studlab = paste(Study),
                comb.fixed = FALSE, comb.random = TRUE,  
                method.tau = "REML",
                hakn = F, prediction = FALSE, sm = "HR")
rema
```

We find there is heterogeneity in OR estimates between studies (Q = 203, 38 df, p<0.0001). Let's see whether there is heterogeneity by study design. We update the meta-analysis object to stratify by study design. 
```{r rema-design}
rema.design <- update.meta(rema, 
                            byvar=Design, 
                            comb.random = TRUE, 
                            comb.fixed = FALSE)
summary(rema.design)
```

```{r forest}
png(c("figures/forest.rema.bydesign.png"), 
    height = 13, width = 6, units='in', res=600)

forest(rema.design,
       
       sortvar = datepub_plot,
       
       rightcols = c("effect.ci", "w.random"),
       rightlabs = c("Ratio Est [95% CI]", "Weight"),
       
       leftcols = c("studlab"),
       leftlab = c("Author, Year"),
       
       just.addcols = "center",
       spacing = 1,
       
       print.tau2 = FALSE,
       print.I2.ci = FALSE,
       
       ref=1,
       col.square = "grey30",
       resid.hetstat=F,
       fontsize = 8)

dev.off()

```

## High Qual Studies

We're going to only keep studies that are prospective cohort studies or population-based case-control studies.

```{r filter-data}
hq.ma.dat = 
  ma.dat %>% 
  filter(study_design %in% c("1. Pro Cht", "2. Pop CC"))
```

Now we repeat the meta-analysis.
```{r ma-hq}
hq.rema <- metagen(lnes, selnes, data = hq.ma.dat, 
                  studlab = paste(Study),
                  comb.fixed = FALSE, comb.random = TRUE,  
                  method.tau = "DL",
                  hakn = F, prediction = FALSE, sm = "HR")
hq.rema
```

Update the meta object to stratify by design 
```{r hq-rema-design}
hq.rema.design <- update.meta(hq.rema, 
                            byvar=Design, 
                            comb.random = TRUE, 
                            comb.fixed = TRUE)
summary(hq.rema.design)
```

```{r forest-bydesign}
png(c("figures/forest.hq.rema.bydesign.png"), 
    height = 13, width = 6, units='in', res=600)

forest(hq.rema.design,
       
       sortvar = datepub_plot,
       
       rightcols = c("effect.ci", "w.random"),
       rightlabs = c("Ratio Est [95% CI]", "Weight"),
       
       leftcols = c("studlab"),
       leftlab = c("Author, Year"),
       
       just.addcols = "center",
       spacing = 1,
       
       print.tau2 = FALSE,
       print.I2.ci = FALSE,
       
       ref=1,
       col.square = "grey30",
       resid.hetstat=F,
       fontsize = 8)

dev.off()

```

## Bias

Create a funnel plot to assess potential for publication bias.

```{r funnel-hhq.rema}
funnel(hq.rema, xlab="Ratio Estimate")
```

This looks pretty suspicious. 

```{r fixed-mbb}
funnel(hq.rema,xlab="Ratio Estimate", 
       contour = c(.95,.975,.99),
       col.contour=c("darkblue","blue","lightblue"))
legend(1.6, 0, 
       c(" 0.05  > p > 0.025",
          "0.025 > p > 0.01", 
               " < 0.01"),
       bty = "n",cex=0.7,
       fill=c("darkblue","blue","lightblue"))
```  

This looks very suspicious. Why?

What does Egger's Test give us?

```{r EggerTest}
et <- metabias(hq.rema, method="mm")
et
```

This shows evidence for publication bias (p=0.001).

Next we'll create a cumulative forest plot.
The cumulative plot is a function in the metafor package, so I have to re-run my meta analysis using the function in the other package.
We will order the studies from lowest standard error (most precision) to highest standard error (least precision).
```{r cumplot}
cuml_meta <- metafor::rma(yi=lnes, sei=selnes, 
                         data=hq.ma.dat, 
              slab=paste(first_auth, datepub, sep=", "))

tmp <- cumul(cuml_meta, order=order(hq.ma.dat$selnes))
 ### cumulative forest plot
forest(tmp, xlim=c(-1,1), at=log(c(0.5, 1, 2)), 
       atransf=exp, digits=c(2,3), cex=0.75)
 
### switch to bold font
par(cex=0.75, font=2)
 
### add column headings to the plot
text(-1, 16, "Author(s) and Year",  pos=4)
text( 1, 16, "Ratio Est [95% CI]", pos=2)
```

Small studies seem to pull the estimate to higher and higher values. If there was no bias, we'd expect the cumulative estimates to fluctuate around a mean.

## Leave one out analysis

```{r metainf}
mi <- metainf(hq.rema, pooled="random", sortvar=hq.ma.dat$datepub)
forest(mi, at = c(0.8,0.9,1.0,1.1,1.2,1.3,1.4,1.5,1.6), 
       xlim=c(0.8,1.6))
```

There is hardly a difference in the summary estimate if you omit any single study.

## Meta-Regression

Let's see if the estimates vary by the proportion of female cases in the study.

```{r data-casesbysex}
#hq.ma.dat %>% select(male_cases,female_cases,cases)

hq.ma.dat <- hq.ma.dat %>%
              mutate(pctfemale = female_cases/cases,
                     pctmale = 1-pctfemale)
```

```{r data-sexratio}
#hq.ma.dat %>% select(lnes,male_cases,female_cases,cases,pctfemale,pctmale)
```



```{r mreg_pctfem}
ma <- metagen(TE=lnes, seTE=selnes, 
        studlab = paste(first_auth, datepub, sep=", "), 
 sm="Ratio Est.", comb.fixed = F, data=hq.ma.dat[!is.na(hq.ma.dat$pctfemale),])
ma.mr <- metareg(ma,pctfemale,method.tau = "REML")
print(ma.mr)
```

```{r bubbleplot}
bubble(ma.mr,ylab="log TE",xlab="Percent Female")
```

What is the estimate in males?

```{r mreg_estmales}
TE.0 <- coef(ma.mr)["intrcpt"]
seTE.0 <- sqrt(vcov(ma.mr)["intrcpt", "intrcpt"])
print(metagen(TE.0, seTE.0, sm="RR"), digits=2)
```

What is the estimate in females?

```{r mreg_pctmale}
ma.mr <- metareg(ma,pctmale,method.tau = "REML")
print(ma.mr)
bubble(ma.mr,ylab="log TE",xlab="Percent Male")

```

```{r mreg_estfemales}
TE.0 <- coef(ma.mr)["intrcpt"]
seTE.0 <- sqrt(vcov(ma.mr)["intrcpt", "intrcpt"])
print(metagen(TE.0, seTE.0, sm="RR"), digits=2)
```


## SessionInfo

```{r si}
sessionInfo()
```